---
title: "Axial-2023-cell-counts"
output: html_document
date: "2024-06-17"
---
Data import and cleanup
```{r}
library(tidyverse)
library(readr)

Axialcellcounts <- read_csv("input-data/Axialcellcounts.csv")
View(Axialcellcounts)

Axial_cellcounts <- read_csv("input-data/Axial_cellcounts - Sheet1.csv")
# View(Axial_cellcounts)

is.na(Axialcellcounts) <- Axialcellcounts == "n/a" #Changes character "n/a" to a true NA

counts <- Axialcellcounts %>%
  filter(!is.na(CELL_ML))
# head(counts)
```

Plot cells counts by location type
```{r}
# ?geom_jitter

library(ggplot2)

str(counts) #check structure of data. CELL_ML is col_character() so must convert to numeric
counts$CELL_ML <- as.numeric(counts$CELL_ML)
```
Make the plot
```{r}
ggplot(counts, aes(x = VENT, y = CELL_ML, color = LOC)) +
  geom_jitter() + 
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(cols=vars(LOC), space = "free", scales = "free") +
  labs(x = "Sample Location", y = "Cells/ml (log)")

```
1. Make the above plot ready to be presented at Lab meeting for Julie. 
```{r}
#Remove uneccesary columns & rename column LOC to LOCATION
refinedcounts <- counts %>% select(-"lOCATION") %>%
  filter( LOC != "background")

names(refinedcounts)[names(refinedcounts) == "LOC"] <- "Location"

# Calculate the average of y
(81900 + 65300) / 2
average_background <- 73600

# Add a horizontal line for the average of y
#plot + geom_hline(yintercept = average_background, color = "red", linetype = "dashed")

#Rename above_vent to niskin
refinedcounts$Location[refinedcounts$Location == "above_vent"] <- "Niskin"
refinedcounts$Location[refinedcounts$Location == "plume"] <- "Plume"
refinedcounts$Location[refinedcounts$Location == "vent"] <- "Vent"

ggplot(refinedcounts, aes(x = VENT, y = CELL_ML, color = Location)) +
  geom_jitter() + 
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(cols=vars(Location), space = "free", scales = "free") +
  labs(x = "Sample Location", y = "Cells/ml (log)") + 
  geom_hline(yintercept = average_background, color = "red", linetype = "dashed")

```

2. Make box plots with the same data. 4 boxes total, for each location categories.
3. Use write.csv to make a CSV file to share

  facet_grid(cols = vars(LOC), space = "free", scales = "free") +
  labs(x = "Sample Location", y = "Cells/ml (log)")

1. Make the above plot ready to be presented at Lab meeting for Julie.
2. Make boxplots with the same data. 4 boxes total, for each of your location categories.
3. Use "write.csv" to make a csv file to share. 

Plot by depth
```{r}
str(Axial_cellcounts)
#Filter by plume samples we want and background
#Plot depth by cell count
#subset samples that have plume and in vent have background
#grepl looks for plume anywhere
counts %>% 
  filter(grepl("Plume", `SAMPLE_ID(s)`) | VENT == "background") %>% 
  mutate(WATER_COLUMN_CATEGORY = case_when(
    DEPTH > 1000 ~ "DEEP",
    DEPTH < 450 ~ "SHALLOW"
  )) %>%
  ggplot(aes(x = CELL_ML, y = DEPTH, fill = `SAMPLE_ID(s)`)) +
  geom_point(shape=21, color = "black") +
    scale_x_log10() +
    scale_y_reverse() +
    facet_grid(rows = vars(WATER_COLUMN_CATEGORY), cols = vars(`SAMPLE_ID(s)`), scales = "free")
  
  
```
  
## Plot by depth

```{r}
str(counts)
# Filter by plume samples we want and the background

# Plot depth by cell count.
counts %>% 
  filter(grepl("Plume", `SAMPLE_ID(s)`) | VENT == "background") %>% 
  mutate(WATER_COLUMN_CATEGORY = case_when(
    DEPTH > 1450 ~ "DEEP",
    DEPTH < 450 ~ "SHALLOW"
  )) %>% 
  ggplot(aes(x = CELL_ML, y = DEPTH, fill = `SAMPLE_ID(s)`)) +
    geom_point(shape = 21, color = "black") +
    scale_x_log10() +
    scale_y_reverse() +
  facet_grid(rows = vars(WATER_COLUMN_CATEGORY), cols = vars(`SAMPLE_ID(s)`), scales = "free")
```