---
title: "microeuk-cell-count-Axial2023"
output: html_document
date: "2024-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

```{r}
library(tidyverse)
library(scales)
```

# Input data

In the shared Axial Cruise 2023 google drive, and under ALL cruise data, there is a sheet called *Grazing-FLP-counts*. Individual sheets from this spreadsheet are used for counting all cells.

Navigate to sheet of interest and export as a .csv from google drive, save in *input-data*

# Eukaryote cell counts

Import raw cell count data

```{r}
# Import csv
counts_raw <- read.csv("input-data/Grazing-FLP-counts - axial23_eukcounts.csv")

# Change "NAs" to a zero.
counts_raw[is.na(counts_raw)] <- 0

# head(counts_raw)
```

## Set up microscopy stats

Determine diameter of FOV, grid, etc.

```
CELLS_PER_ML = ((MEAN)*(Area_chamber))/((Volume)*(Area_FOV)*(Dilution_factor))
```

| Magnification | Diameter of FOV (µm) | Area of FOV (µm\^2) | Area of Grid (µm\^2) | Area of small square in grid (µm\^2) |
|---------------|---------------|---------------|---------------|---------------|
| 10X           | 2200                 | 3801327.11          | 1000000              | 10000                                |
| 20X           | 1100                 | 950331.78           | 250000               | 2500                                 |
| 40X           | 550                  | 237582.94           | 62500                | 625                                  |
| 63X           | 360                  | 101787.60           | 25600                | 256                                  |
| 100X          | 220                  | 38013.27            | 10000                | 100                                  |

```{r}
# Area of chamber we are using for filtering samples
Area_chamber_LgGlass_Hu <- 153.938 #mm^2
Area_chamber_Huber <- 283.5288 #mm^2
Area_chamber_Aceglass_Hu <- 314.1593
# Volume <- 2 #ml # This changes!
# This depends on 
# Area_FOV <- 1E4 #µm^2 at 100X magnification based on how sample is counted.
Dilution_factor <- 0.9 # Pretty standard for what we are doing.

# Make an area of grid key
area_grid <- data.frame(MAG = c(10, 20, 40, 63, 100),
                        area_grid_mm = c(1.00000, 0.24010, 0.06250, 0.02560, 0.01000))
# 1 um = 0.001 mm
```


## Calculate cells per ml

```{r}
counts_processed <- counts_raw %>% 
  # Rename column names
  select(SAMPLE, SAMPLE.ID, EXP.ID, VOL = vol..ml., MAG = Mag..x., count_less20 = X.20um..no.FLP, count_more20 = X.20um..no.FLP.1) %>% 
  # Make a total count column
  mutate(count_total = (count_less20 + count_more20)) %>% 
  # Add in appropriate FOV or Area of Grid, based on magnification
  left_join(area_grid) %>% 
  # Get cell count mean for the field of views counted
  group_by(SAMPLE, SAMPLE.ID, EXP.ID, VOL, MAG, area_grid_mm) %>% 
    summarize(NUM_FOV = n(),
              mean_less20 = sum(count_less20)/NUM_FOV,
              mean_more20 = sum(count_more20)/NUM_FOV,
              mean_total = sum(count_total)/NUM_FOV) %>% 
  group_by(SAMPLE, SAMPLE.ID, EXP.ID) %>% 
  summarise(
    CELLS_PER_ML_less20 = ((mean_less20)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor)),
    CELLS_PER_ML_more20 = ((mean_more20)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor)),
    CELLS_PER_ML_total = ((mean_total)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor))
  ) %>% 
  mutate(CELLS_PER_ML_total_SCI = scientific(CELLS_PER_ML_total, digits = 3))


glimpse(counts_processed)
```
Quick plot
```{r}
counts_processed %>% 
  ggplot(aes(x = SAMPLE, y = CELLS_PER_ML_total)) +
    geom_jitter(stat = "identity", width = 0.1, shape = 21, color = "black", fill = "grey") +
    scale_y_log10() +
  labs(x = "", y = "Log cells/ml", title = "Axial Seamount 2023 eukaryote cell counts") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12))
```

Full table to report
```{r}
counts_processed
```


# Grazing experiment cell counts

```{r}
grazecounts <- read.delim("input-data/Grazing-FLP-counts - axial23_grazingcounts.csv", 
                     blank.lines.skip = FALSE,
                     na.strings = c("", "NA"),
                     stringsAsFactors = FALSE, sep = ",") # Import


grazecounts[is.na(grazecounts)] <- 0 # Change blanks to zeroes

glimpse(grazecounts)
```
## Set microscope information:
```{r}
Area_chamber_LgGlass_Hu <- 153.938 #mm^2

Area_chamber_Huber <- 283.5288 #mm^2

Area_chamber_Aceglass_Hu <- 314.1593

Dilution_factor <- 0.9 # Pretty standard for what we are doing.

# Make an area of grid key (reticule)
area_grid <- data.frame(MAG = c(10, 20, 40, 63, 100),
                        area_grid_mm = c(1.00000, 0.24010, 0.06250, 0.02560, 0.01000))
# 1 um = 0.001 mm
```


## Eukaryote counts from grazing T0


```{r}
cellcounts_from_graze <- grazecounts %>% 
  filter(Notes != "Too difficult to count") %>% 
  select(SAMPLE, SAMPLE.ID, EXP.ID, VOL = vol..ml., MAG = Mag..x., FOV, nanoNoFLP = X.20um..no.FLP, microNoFLP = X.20um..no.FLP.1, nanoFLP = X.20um..w.FLP..comma.sep...cell., microFLP = X.20um..w.FLP..comma.sep...cell..1) %>% 
  # Count number of euk cells observed with FLPs (ex. if "1,2", 'occur' will = 2)
  mutate(nanoFLP_occur = as.numeric(str_count(nanoFLP, "[1-9]\\d*")), 
         microFLP_occur = as.numeric(str_count(microFLP, "[1-9]\\d*")),
         # Add number of euk cells with FLPs to those without for total number of euk cells
         nanoTOTAL = as.numeric(nanoNoFLP) + nanoFLP_occur, 
         microTOTAL = as.numeric(microNoFLP) + microFLP_occur,
         euksTOTAL = nanoTOTAL + microTOTAL)

glimpse(cellcounts_from_graze)
```

```{r}
Area_chamber_LgGlass_Hu <- 153.938 #mm^2

Area_chamber_Huber <- 283.5288 #mm^2

Area_chamber_Aceglass_Hu <- 314.1593

Dilution_factor <- 0.9 # Pretty standard for what we are doing.

# Make an area of grid key (reticule)
area_grid <- data.frame(MAG = c(10, 20, 40, 63, 100),
                        area_grid_mm = c(1.00000, 0.24010, 0.06250, 0.02560, 0.01000))
```


```{r}
counts_cellsml_all <- cellcounts_from_graze %>%
  group_by(SAMPLE, SAMPLE.ID, EXP.ID, VOL = as.numeric(VOL), MAG) %>% #Calculate averages by sample
  summarise(totalFOV = n(), # Count total FOV counted
            nanoAvg = sum(nanoTOTAL)/totalFOV, #Average per FOV
            nanoVar = var(nanoTOTAL), #Variance
            nanoSd = (2*(sqrt(nanoVar))), #Standard deviation
            microAvg = sum(microTOTAL)/totalFOV, ## Repeat for microeuks
            microVar = var(microTOTAL), 
            microSd = (2*(sqrt(microVar))), 
            euksAvg = sum(euksTOTAL)/totalFOV, ## Repeat for total cell count
            euksVar = var(euksTOTAL), 
            euksSd = (2*(sqrt(euksVar))), 
            .groups = 'drop_last') %>%
  left_join(area_grid) %>% 
# Use above microscope stats to fill in:
  mutate(nanoCONC = ((nanoAvg * Area_chamber_LgGlass_Hu)/(VOL * area_grid_mm * 0.9)),
         microCONC = ((microAvg * Area_chamber_LgGlass_Hu)/(VOL * area_grid_mm * 0.9)),
         eukCONC = ((euksAvg * Area_chamber_LgGlass_Hu)/(VOL * area_grid_mm * 0.9))) 

glimpse(counts_cellsml_all)
unique(counts_cellsml_all$SAMPLE.ID)
```

Quick plot for euk cell counts

```{r}
counts_cellsml_all %>% 
  separate(EXP.ID, into = c("REP", "TIMEPOINT"), remove = FALSE) %>% 
  separate(SAMPLE, into = c("SITE", "EXP_tmp"), remove = FALSE) %>% 
  mutate(EXP_TYPE = case_when(
    grepl("miniSID", SAMPLE.ID) ~ "miniSID",
    SITE == "BSW" ~ "large",
    TRUE ~ EXP_tmp)) %>% 
  ggplot(aes(x = SITE, y = eukCONC, fill = TIMEPOINT, shape = REP)) +
  geom_jitter(stat = "identity", width = 0.1, shape = 21, color = "black") +
    scale_y_log10() +
  labs(x = "", y = "Log cells/ml", title = "Eukaryote cell counts from grazing experiments") +
  theme_light() +
  facet_grid(cols = vars(EXP_TYPE), scales = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12))
```


# Grazing experiment calculations

```{r}
glimpse(grazecounts)
unique(grazecounts$Notes)

```

## Base grazing experiment stats
```{r}
graze_0 <- grazecounts %>% 
  filter(Notes != "Too difficult to count" & Notes != "not countable") %>% 
  select(SAMPLE, SAMPLE.ID, EXP.ID, VOL = vol..ml., MAG = Mag..x., FOV, nanoNoFLP = X.20um..no.FLP, microNoFLP = X.20um..no.FLP.1, nanoFLP = X.20um..w.FLP..comma.sep...cell., microFLP = X.20um..w.FLP..comma.sep...cell..1) %>%
  separate(EXP.ID, into = c("REP", "TIMEPOINT"), remove = FALSE) %>% 
  separate(SAMPLE, into = c("SITE", "EXP_tmp"), remove = FALSE) %>% 
  mutate(EXP_TYPE = case_when(
    grepl("miniSID", SAMPLE.ID) ~ "miniSID",
    SITE == "BSW" ~ "large",
    TRUE ~ EXP_tmp)) 

glimpse(graze_0)
```
Total number of grazing experimentss
```{r}
# Sites estimated
unique(graze_0$SITE)

# Types of experiments considered
unique(graze_0$EXP_TYPE)
```


## Pull out ingested flps
```{r}
graze_euksw_flp <- graze_0 %>% 
  select(-nanoNoFLP, -microNoFLP) %>% 
  # Inputs that are comma separated will be split into a new row
  separate_rows(microFLP, sep = ",", convert = TRUE) %>%
  separate_rows(nanoFLP, sep = ",", convert = TRUE) %>%
  # Replace NAs with zeroes
  replace_na(list(microFLP = 0, nanoFLP = 0)) %>% 
  select(micro_size = microFLP, nano_size = nanoFLP, everything()) %>% 
  pivot_longer(cols = ends_with("_size"), names_to = "SizeFrac", values_to = "num_of_FLP") %>%
  filter(num_of_FLP > 0) %>% 
  group_by(SAMPLE.ID, EXP.ID, REP, TIMEPOINT, EXP_TYPE, SizeFrac, SAMPLE, SITE) %>% 
    summarise(total_FLP = sum(num_of_FLP),
              total_euks_wflp = n())


glimpse(graze_euksw_flp)
```
1. total_FLP = sum of FLPs found inside a euk cell

2. total_euks_wflp = number of euks counted with ingested FLP

> Repeat above operation for euk cells without any FLP. Here, subset total number of observations where there was a euk cell without FLP. These need to be counted as euk cell without an FLP.

## Estimate euks with out FLPs

```{r}
glimpse(graze_0)
```

This is repeated because we need to count the number of euk cells without FLP in the overall average.
```{r}
graze_1 <- graze_0 %>% 
  # Select cells without FLPs to count.
  select(-nanoFLP, -microFLP) %>% 
  select(micro_size = microNoFLP, nano_size = nanoNoFLP, everything()) %>% 
  pivot_longer(cols = ends_with("_size"), names_to = "SizeFrac", values_to = "num_of_euks") %>%
  group_by(SAMPLE.ID, EXP.ID, REP, TIMEPOINT, EXP_TYPE, SizeFrac, SAMPLE, SITE) %>% 
    summarise(total_euks_noFLP = sum(num_of_euks)) %>% 
  #
  # Join with FLP count dataframe
  left_join(graze_euksw_flp) %>% 
  replace_na(list(total_FLP = 0, total_euks_wflp = 0)) #Replace NAs with zero

glimpse(graze_1)
unique(graze_1$TIMEPOINT)
```

Compile all grazing, full table:
```{r}
counts_grazing_compiled <- graze_1 %>% 
  group_by(SITE, SAMPLE, SAMPLE.ID, EXP.ID, EXP_TYPE, REP, TIMEPOINT) %>% 
  summarise(total_euks_noFLP = sum(total_euks_noFLP),
            total_FLP = sum(total_FLP),
            total_euks_wFLP = sum(total_euks_wflp)) %>% 
  mutate(MINUTES = as.numeric(str_remove(TIMEPOINT, "T"))) %>% 
  group_by(SITE, SAMPLE, SAMPLE.ID, EXP.ID, EXP_TYPE, REP, TIMEPOINT, MINUTES) %>% 
  #
  # FLPperEuk is the total FLP divided by the total number of euk cells counted
  mutate(FLPperEuk = total_FLP/(sum(total_euks_noFLP, total_euks_wFLP)))

  
glimpse(counts_grazing_compiled)
# If we wanted to estimate this by size fraction, this is where we would integrate total with size fractionated.
```

The code above takes the processed microscopy counts, and estimates the total number of FLP and the total number of eukaryotes counted (with or without FLP inside). To calculate the rate that FLP were ingested, we need the slope of the line when we plot the number of ingested FLP per eukaryote cell by experiment time.

We will use this equation below: `FLPperEuk = total_FLP/(sum(total_euks_noFLP, total_euks_wflp))`.

### Plot FLP per euk by time

```{r, fig.width=9, fig.height=2}
counts_grazing_compiled %>% 
  ggplot(aes(x = MINUTES, y = FLPperEuk, fill = EXP_TYPE)) +
  geom_point(stat = "identity", shape = 21, color = "black") +
    facet_grid(cols = vars(SITE)) +
  theme_linedraw()
```

```{r}
glimpse(counts_grazing_compiled)
```


## Run lm to get slope

```{r}
library(broom)
```


```{r}

calculate_lm <- function(df){
  regression_1 <- df %>%
  type.convert(as.is = TRUE) %>%
  #
  group_by(SITE, SAMPLE, SAMPLE.ID, EXP_TYPE) %>%
    nest(-SITE, -SAMPLE, -SAMPLE.ID, -EXP_TYPE) %>% 
  #
  mutate(lm_fit = map(data, ~lm(FLPperEuk ~ MINUTES, data = .)),
         tidied = map(lm_fit, tidy)) %>% 
  unnest(tidied) %>% 
  select(SITE, SAMPLE, SAMPLE.ID, EXP_TYPE, term, estimate) %>%
  #
  pivot_wider(names_from = term, values_from = estimate) %>% 
  data.frame
  # Reset column names
  colnames(regression_1) <- c("SITE", "SAMPLE", 
                              "SAMPLE.ID", "EXP_TYPE",
                              "INTERCEPT", "SLOPE")
  # Repeat broom model to get R2
  out_regression <- df %>%
  group_by(SITE, SAMPLE, SAMPLE.ID, EXP_TYPE) %>%
    nest(-SITE, -SAMPLE, -SAMPLE.ID, -EXP_TYPE) %>% 
  #
  mutate(lm_fit = map(data, ~lm(FLPperEuk ~ MINUTES, data = .)),
         glanced = map(lm_fit, glance)) %>% 
  unnest(glanced) %>% 
    select(SITE, SAMPLE, SAMPLE.ID, EXP_TYPE, r.squared) %>%
  right_join(regression_1) %>% 
  right_join(df) %>% 
  data.frame
  return(out_regression)
}
```

```{r}
grazing_compiled_wslope <- calculate_lm(counts_grazing_compiled)
#
glimpse(grazing_compiled_wslope)
# unique(grazing_compiled_wslope$SLOPE)
```

