---
title: "microeuk-cell-count-Axial2023"
output: html_document
date: "2024-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

```{r}
library(tidyverse)
library(scales)
```

# Input data

In the shared Axial Cruise 2023 google drive, and under ALL cruise data, there is a sheet called *Grazing-FLP-counts*. Individual sheets from this spreadsheet are used for counting all cells.

Navigate to sheet of interest and export as a .csv from google drive, save in *input-data*

# Eukaryote cell counts

Import raw cell count data

```{r}
# Import csv
counts_raw <- read.csv("input-data/Grazing-FLP-counts - axial23_eukcounts.csv")

# Change "NAs" to a zero.
counts_raw[is.na(counts_raw)] <- 0

# head(counts_raw)
```

## Set up cells per ml calc

Determine diameter of FOV, grid, etc.

```
CELLS_PER_ML = ((MEAN)*(Area_chamber))/((Volume)*(Area_FOV)*(Dilution_factor))
```

| Magnification | Diameter of FOV (µm) | Area of FOV (µm\^2) | Area of Grid (µm\^2) | Area of small square in grid (µm\^2) |
|---------------|---------------|---------------|---------------|---------------|
| 10X           | 2200                 | 3801327.11          | 1000000              | 10000                                |
| 20X           | 1100                 | 950331.78           | 250000               | 2500                                 |
| 40X           | 550                  | 237582.94           | 62500                | 625                                  |
| 63X           | 360                  | 101787.60           | 25600                | 256                                  |
| 100X          | 220                  | 38013.27            | 10000                | 100                                  |

```{r}
# Area of chamber we are using for filtering samples
Area_chamber_LgGlass_Hu <- 153.938 #mm^2
Area_chamber_Huber <- 283.5288 #mm^2
Area_chamber_Aceglass_Hu <- 314.1593
# Volume <- 2 #ml # This changes!
# This depends on 
# Area_FOV <- 1E4 #µm^2 at 100X magnification based on how sample is counted.
Dilution_factor <- 0.9 # Pretty standard for what we are doing.

# Make an area of grid key
area_grid <- data.frame(MAG = c(10, 20, 40, 63, 100),
                        area_grid_mm = c(1.00000, 0.24010, 0.06250, 0.02560, 0.01000))
# 1 um = 0.001 mm
```

```{r}
counts_processed <- counts_raw %>% 
  # Rename column names
  select(SAMPLE, SAMPLE.ID, EXP.ID, VOL = vol..ml., MAG = Mag..x., count_less20 = X.20um..no.FLP, count_more20 = X.20um..no.FLP.1) %>% 
  # Make a total count column
  mutate(count_total = (count_less20 + count_more20)) %>% 
  # Add in appropriate FOV or Area of Grid, based on magnification
  left_join(area_grid) %>% 
  # Get cell count mean for the field of views counted
  group_by(SAMPLE, SAMPLE.ID, EXP.ID, VOL, MAG, area_grid_mm) %>% 
    summarize(NUM_FOV = n(),
              mean_less20 = sum(count_less20)/NUM_FOV,
              mean_more20 = sum(count_more20)/NUM_FOV,
              mean_total = sum(count_total)/NUM_FOV) %>% 
  group_by(SAMPLE, SAMPLE.ID, EXP.ID) %>% 
  summarise(
    CELLS_PER_ML_less20 = ((mean_less20)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor)),
    CELLS_PER_ML_more20 = ((mean_more20)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor)),
    CELLS_PER_ML_total = ((mean_total)*(Area_chamber_LgGlass_Hu))/((VOL)*(area_grid_mm)*(Dilution_factor))
  ) %>% 
  mutate(CELLS_PER_ML_total_SCI = scientific(CELLS_PER_ML_total, digits = 3))


glimpse(counts_processed)
```
Quick plot
```{r}
counts_processed %>% 
  ggplot(aes(x = SAMPLE, y = CELLS_PER_ML_total)) +
    geom_jitter(stat = "identity", with = 0.2) +
    scale_y_log10() +
  labs(x = "", y = "Log cells/ml", title = "Axial Seamount 2023 eukaryote cell counts") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12))
```

Full table to report
```{r}
counts_processed
```


```{r}
euk_cell_conc <- counts_raw %>%
    group_by(SAMPLE, SAMPLE.ID, EXP.ID) %>%
    summarise(Avg_cellmL = mean(Cellsperml), # Average cells per ml across replicates
              sem=sd(Cellsperml)/sqrt(length(Cellsperml)), # Standard mean error
              SD=sd(Cellsperml),  #standard deviation
              var=sqrt(SD),  # variance
              Num = n()) %>% #Total number of 
    data.frame
```
